---
title: "Viz Template"
author: "Dr. Chris J. Martin"
date: "`r format(Sys.time(), '%a %d %B, %Y')`"
output: html_document 
editor_options: 
  markdown: 
    wrap: 72
---

## Set up

```{r, setup}
knitr::opts_chunk$set(echo = TRUE)

# this is for rendering custom fonts in markdown
# but it doesn;t appear to be working ...

ragg_png_wrapper = function(..., res = 150) {
  ragg::agg_png(..., res = res, units = "in")
}

knitr::opts_chunk$set(dev = ragg_png_wrapper)
```

```{r}
library(tidyverse)

library(ggtext)     # improved text rendering for ggplot2
library(ggforce)    # extra functionality for ggplot2 inc. improved annotation
library(patchwork)  # composing layouts with multiple plots
library(ggdist)     # visualising uncertainty and distributions
library(readxl)     # for reading Excel files
library(lubridate)  # for dates
library(readxl)     # for reading excel files
library(patchwork)  # for composing layouts of multiple plots

# packages for using custom fonts
library(systemfonts)
library(textshaping)

library(cjmr)       # my own package including default styling for plots

# set default theme for exploratory plots
theme_set(theme_light())
```

## Source of info on common ggplot2 tasks

-   A short and general reference to common tasks at the BBC:
    <https://bbc.github.io/rcookbook/>

-   ggplot2 wizardry by Cédric Scherer -
    <https://z3tt.github.io/OutlierConf2021/>

-   R Graphics Cookbook by W. Chang - <https://r-graphics.org/>

-   Custom fonts -
    <https://yjunechoe.github.io/posts/2021-06-24-setting-up-and-debugging-custom-fonts/>

## Useful code snippets

[Changing the size of a
legend](https://z3tt.github.io/OutlierConf2021/#legend-design) - Cédric
Scherer

```{r, eval=FALSE}
guides(color = guide_colorbar(title.position = "top", 
                                title.hjust = .5, 
                                barwidth = unit(20, "lines"), 
                                barheight = unit(.5, "lines")))

```

[Changing margins](https://z3tt.github.io/OutlierConf2021/#white-space)
- Cédric Scherer

```{r, eval=FALSE}
theme(plot.margin = margin(t = 25, r = 25, b = 10, l = 25)) # top, right, bottom, left
```

Adjusting axis

```{r, eval=FALSE}
scale_y_continuous(breaks = seq(0,350, 50), # labels
                   limits = c(0, 350)) # extent of axis
```

## Colour palettes

## Remember it is an iterative process

## Import data

```{r}
raw_data <- haven::read_dta("migrazensus_v0.1.dta")
```

## Clean data

Identifiers

• wknr_2021: Official ID number for electoral district

• wkname_2021: Name of electoral district

• land: Name of federal state

• landid: Official numeric ID for federal state

Estimates obtained from micro census

• bevoelkerung: Estimated population size for a district

• wahlberechtigte: Estimated number of eligible voters

• wbmighintergrund_absolut: Number of eligible citizen with migrant
background

• wbmighintergrund_anteil: Share of eligible citizens with migrant
background among all eligible voters

• mighintergrund_absolut: Number of residents with migrant background

• mighintergrund_anteil: Share of residents with migrant background
among all residents (population)

2017 results within 2021 boundaries

1

• wahlberechtigte_2017: Number of eligible voters

• waehler: Number of voters

• gueltige_1: Number of valid votes (Erststimme, i.e., vote for a
district candidate)

• gueltige_2: Number of valid votes (Zweitstimme, i.e., vote for a
district candidate)

• cdu_1: Number of votes for CDU (Erststimme)

• cdu_2: Number of votes for CDU (Zweitstimme)

• spd_1: Number of votes for SPD (Erststimme)

• spd_2: Number of votes for SPD (Zweitstimme)

• linke_1: Number of votes for Die Linke (Erststimme)

• linke_2: Number of votes for Die Linke (Zweitstimme)

• gruene_1: Number of votes for Bündnis 90/Die Grünen(Erststimme)

• gruene_2: Number of votes for Bündnis 90/Die Grünen(Zweitstimme)

• csu_1: Number of votes for CSU (Erststimme)

• csu_2: Number of votes for CSU (Zweitstimme)

• fdp_1: Number of votes for FDP (Erststimme)

• fdp_2: Number of votes for FDP (Zweitstimme)

• afd_1: Number of votes for AfD (Erststimme)

• afd_2: Number of votes for AfD (Zweitstimme)

Note: Turnout can be calculated as waehler / wahlberechtigte_2017 and
vote shares as, for instance,

cdu_1 / gueltige_1 oder spd_2 / gueltige_2.

```{r}
raw_data %>% 
  visdat::vis_miss()

clean_data <- raw_data %>% 
  select(district_id = wknr_2021,
         district_nm = wkname_2021,
         mg_bg = mighintergrund_absolut, 
         elig_voters_all_1 = wahlberechtigte, 
         elig_voters_all_2 = waehler,
         elig_voters_mg_bg = wbmighintergrund_absolut,
         perc_elig_voters_mg_bg = wbmighintergrund_anteil,
         ends_with("1"),
         ends_with("2"))

clean_data %>% 
  View()
```

## Explore the data

1.  Show that eligible citizens with migration background matter for who
    wins the constituency ("Erststimme") -\> e.g., my table above, your
    ideas?

```{r}
# code from Arndt Leininger 
winner_runner_up <- raw_data %>% 
  select(wknr_2021, contains('_1'), -gueltige_1, -alq_15_24) %>% 
    pivot_longer(cols = cdu_1:afd_1, names_to = 'party', values_to = 'votes') %>% 
    group_by(wknr_2021) %>% 
    arrange(wknr_2021, desc(votes)) %>% 
    slice(1:2) %>% 
    summarize(first_placed = first(party),
              first_placed_votes = first(votes),
              second_placed = nth(party, 2),
              second_placed_votes = nth(votes, 2),
              margin = first_placed_votes - second_placed_votes)

raw_data <- left_join(raw_data, winner_runner_up, by = 'wknr_2021')

swing_districts <- raw_data %>% 
  arrange(margin) %>% 
  mutate(turnout = waehler / wahlberechtigte_2017) %>% 
  select(state = land,
         state_id = landid,
         district_id = wknr_2021,
         district_nm = wkname_2021,
         first_placed, first_placed_votes, 
         second_placed, second_placed_votes,
         margin,
         elig_voters_mg_bg = wbmighintergrund_absolut,
         turnout) %>% 
  mutate(non_voting_mg_bg = elig_voters_mg_bg * (1 - turnout),
         potential_to_swing_election = non_voting_mg_bg > margin)

swing_districts
```

## Explore the context of the data

-   **Who collected the data?**

-   **Why was the data collected?**

-   **What is the data used for?**

-   **Are there some other datasets that might be useful for
    comparisons?**

## Think about the audience

+----------------------------------------------------------------+-----+
| To think about                                                 | No  |
|                                                                | tes |
+================================================================+=====+
| Can you think of a group of people who might view the          |     |
| visualisations?                                                |     |
+----------------------------------------------------------------+-----+
| Can you think of a specific person who might view the          |     |
| visualizations?                                                |     |
+----------------------------------------------------------------+-----+
| What do want the audience to know or be able to after viewing  |     |
| the visualizations?                                            |     |
+----------------------------------------------------------------+-----+
| What expectations might the audience have about visualisations |     |
| of this data?                                                  |     |
+----------------------------------------------------------------+-----+

## Ask some questions to be answered by the visualizations

1.  

2.  Show that eligible citizens with migration background matter for the
    party vote (seats are allocated at the Laender level) -\> your
    ideas?

3.  Show that non-enfranchised residents with migration background
    (underage citizens with migration background and alien residents)
    could matter, too, if naturalized or enfranchised. -\> e.g., my
    parliament plot, your ideas?

## Summarize the story that the visualizations can help tell

Based on the questions above, the story might go something like this:

-   ...

-   ...

## Create the visualisation/s

### Visualization 1 (question n)

+---------------------+-----------------+---------------------+
| Question            | Answer (i.e.    | Notes to consider   |
|                     | title)          | in design, chart    |
|                     |                 | types etc.          |
+=====================+=================+=====================+
| 1.(a) How many      |                 |                     |
| people with migrant |                 |                     |
| backgrounds are     |                 |                     |
| there living in     |                 |                     |
| Germany?            |                 |                     |
+---------------------+-----------------+---------------------+
| 1.(b) What          |                 |                     |
| proportion of       |                 |                     |
| people with a       |                 |                     |
| migrant background  |                 |                     |
| are eligible to     |                 |                     |
| vote?               |                 |                     |
+---------------------+-----------------+---------------------+
| 1.  (c) How many    |                 |                     |
|     seats is this   |                 |                     |
|     equivalent to?  |                 |                     |
+---------------------+-----------------+---------------------+
| 1.(d) Where do      |                 |                     |
| people with a       |                 |                     |
| migrant background  |                 |                     |
| who are eligible to |                 |                     |
| vote live?          |                 |                     |
+---------------------+-----------------+---------------------+

```{r}
q1_plotting_df <- raw_data %>% 
  select(district_id = wknr_2021,
         district_nm = wkname_2021,
         state_nm = land,
         state_id = landid,
         mg_bg = mighintergrund_absolut, 
         elig_voters_all_1 = wahlberechtigte, 
         elig_voters_all_2 = waehler,
         elig_voters_mg_bg = wbmighintergrund_absolut) %>% 
  mutate(across(where(is.numeric), round)) %>% 
  group_by(state_nm) %>% 
  summarise(mg_bg = sum(mg_bg),
            elig_voters_mg_bg = sum(elig_voters_mg_bg)
            ) %>% 
  mutate(prop_eligible = elig_voters_mg_bg / mg_bg,
         prop_inelgible = 1 - prop_eligible)


  
q1a_df <- q1_plotting_df %>% 
  mutate(highlight = mg_bg > 1e+06)

total_mg_bg <- sum(q1a_df$mg_bg)

mg_bg_highlighted_states <- q1a_df %>% 
  filter(highlight) %>% 
  summarise(sum(mg_bg)) %>% 
  pull()

prop_mg_bg_in_hl_states <- mg_bg_highlighted_states / total_mg_bg

annotation <- glue::glue("Approximately {round(total_mg_bg / 1e+06, 1)} million",
                         " people",
                         "\nwith a migrant background live in Germany.\n",
                         "{round(prop_mg_bg_in_hl_states * 100)}% of these people",
                         " live in the six highlighted states")

q1a <- q1a_df %>% 
  ggplot(aes(mg_bg / 1e+06, 
             fct_reorder(state_nm, mg_bg),
             fill = highlight)) +
  geom_col() +
  annotate("text",
           x = 3, y = "Bremen",
           label = annotation, family = "Lato") +
  labs(x = "Millions of people with a migrant background") +
  
  theme_cjmr_explanatory() +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line(),
        legend.position = "none")

q1a

ggsave("q1a.svg", plot = q1a, units = "mm",
       width = 140, height = 90)
  
```

```{r}
q1b_df <- q1a_df %>% 
  select(state_nm, prop_eligible, prop_inelgible, mg_bg) %>% 
  pivot_longer(c(prop_eligible, prop_inelgible), 
               names_to = "prop", values_to = "num")


q1b <- q1b_df %>% 
  filter(prop == "prop_eligible") %>% 
  mutate(highlight = num > 0.4) %>% 
  ggplot(aes(num, fct_reorder(state_nm, mg_bg), 
             fill = highlight)) +
  geom_vline(xintercept = 0.4) +
  geom_col() +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  
  theme_cjmr_explanatory() +
    theme(panel.grid.major.y = element_blank(),
          panel.grid.major.x = element_line(),
        legend.position = "none") +
  
  labs(x = "Percentage of people with a migrant background who are eligible to vote")

q1b

ggsave("q1b.svg", plot = q1b, units = "mm",
       width = 96.7, height = 90)

```

xdf

```{r}
raw_data %>% 
  select(district_id = wknr_2021,
         district_nm = wkname_2021,
         state_nm = land,
         state_id = landid,
         mg_bg = mighintergrund_absolut, 
         elig_voters_all_1 = wahlberechtigte, 
         elig_voters_all_2 = waehler,
         elig_voters_mg_bg = wbmighintergrund_absolut) %>% 
  summarise(total_voters = sum(elig_voters_all_1),
            total_mg_bg = sum(elig_voters_mg_bg)) %>% 
  mutate()

 

votes_per_seat <- sum(raw_data$waehler) / 598
seats_for_mg_bg_votes <- round(sum(raw_data$wbmighintergrund_absolut) /
                                 votes_per_seat)
seats_for_non_mg_bg_votes <- 598 - seats_for_mg_bg_votes

seats <- tribble(
  ~"bg", ~"seats",
  "mg", seats_for_mg_bg_votes,
  "non_mg", seats_for_non_mg_bg_votes
)

library(waffle)

q1c <- ggplot(seats,
       aes(values = seats, fill = bg)) +
  geom_waffle(colour = "white") + #, flip = TRUE
  coord_equal() +
  theme_void() +
  theme_enhance_waffle() +
  theme(legend.position = "none")

q1c

ggsave("q1c.svg", plot = q1c, units = "mm",
       width = 256.7, height = 30)
```

Working out the majority

Coalition partners seats -
(<https://en.wikipedia.org/wiki/List_of_political_parties_in_Germany>)

-   CDU/CSU - 245

-   SDP - 152

-   total = 397

-   Number of seats total - 709
    (<https://en.wikipedia.org/wiki/Bundestag>)

-   Other - 312

-   Majority - 397-312 = 85

-   

### Visualization 2 (question 2)

+-------------------------+-----------------+-------------------------+
| Question                | Answer (i.e.    | Notes to consider in    |
|                         | title)          | design, chart types     |
|                         |                 | etc.                    |
+=========================+=================+=========================+
| Where can voters with   |                 |                         |
| migration backgrounds   |                 |                         |
| swing elections of      |                 |                         |
| district                |                 |                         |
| representative?         |                 |                         |
+-------------------------+-----------------+-------------------------+

```{r}
library(sf)

districts_geometry <- read_sf("btw21_geometrie_wahlkreise_shp/Geometrie_Wahlkreise_20DBT.shp") %>% 
  janitor::clean_names() %>% 
  rename(district_id = wkr_nr)
  

q2_df <- swing_districts %>% 
  
  # add geometry
  mutate(district_id = as.numeric(district_id)) %>% 
  left_join(districts_geometry) %>% 
  
  # identify districts where migration background votes could swing election
  mutate(pot_to_swing_2 = case_when(
    non_voting_mg_bg > margin ~ "outright",
    non_voting_mg_bg + 1000 > margin ~ "+1000",
    non_voting_mg_bg + 2500 > margin ~ "+2500",
    TRUE ~ "no"),
    pot_to_swing_2 = na_if(pot_to_swing_2, "no"),
    pot_to_swing_2 = factor(pot_to_swing_2,
                            levels = c("outright", "+1000", "+2500"))
  
  ) %>% 
  
  # convert to sf object for plotting
  st_as_sf() %>%
  
  # focus down on useful variables
  select(second_placed, pot_to_swing_2, district_id, district_nm) %>% 
  
  # remove nas 
  filter(!is.na(pot_to_swing_2))
  
 
q2_df %>% 
  mutate(plus_2500 = !is.na(pot_to_swing_2),
         plus_1000 = (pot_to_swing_2 == "+1000" | pot_to_swing_2 == "outright"),
         outright = (pot_to_swing_2 == "outright")) %>% 
  group_by(second_placed) %>% 
  summarise(outright = sum(outright),
            plus_1000 = sum(plus_1000),
            plus_2500 = sum(plus_2500)) %>% 
  filter(outright > 0) %>% 
  
  st_drop_geometry() %>% 
  ggplot(aes(outright, fct_reorder(second_placed, outright),
             fill = second_placed)) +
  geom_col() +
  scale_y_discrete(labels = c(cdu_1 = "CDU", spd_1 = "SPD", 
                              linke_1 = "Linke", gruene_1 = "Gruene")) +
  scale_fill_manual(values = c(cdu_1 = "#666666", spd_1 = "#C20A1B", 
                              linke_1 = "#C20A83", gruene_1 = "#64C30A")) +
  
  labs(x = "Number of district seats that could be won by gaining the votes of
       people with a migrant vote who didn't vote at the last election") +
  
  theme_cjmr_explanatory() +
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.major.x = element_line())
  
  
  ggplot() +
  geom_sf(mapping = aes(alpha = fct_rev(pot_to_swing_2)), fill = "blue") +
  facet_wrap(~ second_placed)
  

  group_by(second_placed) %>% 
  nest()

  # focus on the two main parties (cdu_1 and sdp_1)
  filter(second_placed == "cdu_1" | second_placed == "sdp_1" )
  
  #  
  # 
  # ggplot() + geom_sf()
```

### Visualization x (question n)

+-----------+-----------------+----------------------------------------+
| Question  | Answer (i.e.    | Notes to consider in design, chart     |
|           | title)          | types etc.                             |
+===========+=================+========================================+
|           |                 |                                        |
+-----------+-----------------+----------------------------------------+

```{r}

```

## Export the visualizations for editing

Some useful image sizes (mm unless stated)

+---------------------------------------------------+--------+---------+
| Purpose                                           | Width  | Height  |
+===================================================+========+=========+
| Single plot for Twitter                           |        |         |
+---------------------------------------------------+--------+---------+
|                                                   |        |         |
+---------------------------------------------------+--------+---------+
|                                                   |        |         |
+---------------------------------------------------+--------+---------+

```{r}

file_folder <- ""
file_name <- "holding_name"

ggsave(str_c(file_folder, file_name, ".svg"))

# if saving as a pdf use device = cairo_pdf in ggsave 
```

## Import back in the vizualisation as an image

## Consider how well the overall design works

Based on [Flowing
Data](https://flowingdata.com/visualization-for-clarity/designing-for-an-audience/adjusting-for-the-differences/).

Think about each issue (in bold below) as a continuum rather than a
check-box.

+-----------------------------+---------------------------------+-----+
| Question to consider        | Tips                            | No  |
|                             |                                 | tes |
+=============================+=================================+=====+
| Are the **encodings**       | Air on the side of too much     |     |
| explained?                  | explanation ...                 |     |
+-----------------------------+---------------------------------+-----+
| Is the **context**          | Titles, subtitles               |     |
| explained?                  |                                 |     |
|                             | Annotate                        |     |
|                             |                                 |     |
|                             | Use colour to highlight         |     |
+-----------------------------+---------------------------------+-----+
| Does it have                | Think visual hierarchy *(as     |     |
| **readability** like text?  | when laying out text)*          |     |
+-----------------------------+---------------------------------+-----+
| Do the **aesthetics** align | Serious topic -\> dark colours, |     |
| with the topic/message?     | straight-forward chart types    |     |
+-----------------------------+---------------------------------+-----+

## Remember it is an iterative process
